---
title: "Soil Chemistry: pH Distribution"
author: "McKinley Martin"
date: "2026-07-15"
output: pdf_document
---

This R Markdown file details creating a distribution plot using ggplot2 from a csv file. In particular, this script focuses on the environmental pH distribution of various soil samples. For more information on basic density charts with ggplot2, reference the following link: https://r-graph-gallery.com/21-distribution-plot-using-ggplot2.html \

Start by loading in the ggplot2 package. 
```{r}
library(ggplot2)
```
\
Next, read in the csv file as the function. The head() function allows one to preview the dataset without loading in the entire file. 
```{r, echo=TRUE, warning=FALSE, message=FALSE}
density_function <- read.csv("CIB_Soil_Chemistry_pH.csv")
head(density_function)
```
\
Finally, set up the script for running the plot. In this case, vertical lines were introduced to demonstrate experimental media pH values for evaluation against initial environmental conditions. 
```{r}
ggplot(density_function, aes(x = pH)) + 
  geom_density(fill = "lightblue", alpha = 0.5) + 
  
  # create vertical lines
  geom_vline(xintercept = 5, linetype = "dashed", color = "red") + 
  geom_vline(xintercept = 7, linetype = "dashed", color = "blue") + 
  geom_vline(xintercept = 9, linetype = "dashed", color = "darkgreen") + 
  
  theme_minimal() + 
  
  # create labels
  labs(title = "Distribution of Soil pH Values", x = "Soil pH", y = "Density"
  )
```

