---
title: 'Soil Chemistry: Salinity Distribution'
author: "McKinley Martin"
date: "2026-07-15"
output: pdf_document
---

This R Markdown file details creating a distribution plot using ggplot2 from a csv file. In particular, this script focuses on the environmental salinity distribution of various soil samples. For more information on basic density charts with ggplot2, reference the following link:
\
https://r-graph-gallery.com/21-distribution-plot-using-ggplot2.html \

The following two web pages were referenced in conversion of salinity percentages to dS/m:
\
https://www.landscape.sa.gov.au/mr/publications/measuring-salinity 
\
https://watereuse.org/salinity-management/ls/ls_3d.html \

Start by loading in the ggplot2 package. 
```{r}
library(ggplot2)
```
\
Next, read in the csv file as the function. The as.numeric() function allows one to convert strings to numerical values for plotting. 
```{r, echo=TRUE, warning=FALSE, message=FALSE}
density_function <- read.csv("CIB_Soil_Chemistry_Salinity.csv")
density_function$Salinity_dS_m <- as.numeric(density_function$Salinity_dS_m)
```
\
Finally, set up the script for running the plot. In this case, vertical lines were introduced to demonstrate experimental media salinity values for evaluation against initial environmental conditions. 
```{r, warning=FALSE, message=FALSE}
ggplot(density_function, aes(x = Salinity_dS_m)) + 
  geom_density(fill = "lightblue", alpha = 0.5) + 
  
  # create vertical lines
  geom_vline(xintercept = 0, linetype = "dashed", color = "red") + 
  geom_vline(xintercept = 23.4375, linetype = "dashed", color = "blue") + 
  geom_vline(xintercept = 46.875, linetype = "dashed", color = "darkgreen") + 
  
  theme_minimal() + 
  
  # create labels
  labs(title = "Distribution of Soil Salinity Values", x = "Soil Salinity (dS/m)", 
       y = "Density"
  )
```
